---
title: "Customer Purchase Analysis: R EDA"
author: "Nikita"
output: html_notebook
---

## 1. Introduction

This notebook performs data cleaning and (EDA)
on customer purchase data to understand purchasing behavior and patterns.
```{r}
## 2. Load Libraries

library(tidyverse)
library(readxl)
library(janitor)

```
```{r}
## 3. Load Dataset

df <- read_excel(
  "C:/Users/sniks/OneDrive/Desktop/customer-purchase-analysis-and-dashboard/bike_buyers_analysis.xlsx",
  sheet = "Raw_Data"
)

head(df)
```
```{r}
## 4. Clean Column Names
df <- df %>%
  clean_names()

glimpse(df)
```
```{r}
## 5. Dataset Overview

dim(df)
summary(df)
```
```{r}
## 6. Purchase Distribution

ggplot(df, aes(x = purchased_bike)) +
  geom_bar(fill = "steelblue") +
  labs(
    title = "Purchase Distribution",
    x = "Purchased Bike",
    y = "Count"
  )
```
```{r}
## 7. Purchase by Gender

ggplot(df, aes(x = gender, fill = purchased_bike)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Purchase by Gender",
    x = "Gender",
    y = "Count"
  )
```
```{r}
## 8. Age Distribution

ggplot(df, aes(x = age)) +
  geom_histogram(bins = 20, fill = "darkgreen") +
  labs(
    title = "Age Distribution",
    x = "Age",
    y = "Count"
  )
```
```{r}
## 9. Age Group Creation

df <- df %>%
  mutate(
    age_group = case_when(
      age < 25 ~ "Under 25",
      age >= 25 & age <= 35 ~ "25–35",
      age > 35 & age <= 50 ~ "36–50",
      TRUE ~ "50+"
    )
  )
```
```{r}
## 10. Purchase by Age Group

ggplot(df, aes(x = age_group, fill = purchased_bike)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Purchase by Age Group",
    x = "Age Group",
    y = "Count"
  )
```
```{r}
## 11. Income vs Purchase

ggplot(df, aes(x = purchased_bike, y = income, fill = purchased_bike)) +
  geom_boxplot() +
  labs(
    title = "Income vs Purchase Decision",
    x = "Purchased Bike",
    y = "Income"
  )
```
```{r}
## 12. Commute Distance vs Purchase

ggplot(df, aes(x = commute_distance, fill = purchased_bike)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Commute Distance vs Purchase",
    x = "Commute Distance",
    y = "Count"
  )
```
```{r}
## 13. Region-wise Purchase

ggplot(df, aes(x = region, fill = purchased_bike)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Region-wise Purchase",
    x = "Region",
    y = "Count"
  )
```
```{r}
## 14. Key Metrics

total_customers <- nrow(df)

buyers <- df %>%
  filter(purchased_bike == "Yes") %>%
  nrow()

purchase_rate <- buyers / total_customers

total_customers
buyers
round(purchase_rate * 100, 2)
```
## 15. Key Insights

- Customers with higher income are more likely to purchase.
- Customers aged 30–50 represent the primary buyer segment.
- Shorter commute distances are associated with higher purchase likelihood.
- Purchasing behavior varies significantly across regions.
